---
title: 'Repdata-033: Peer Assessment 1'
author: "AAGP80"
date: "12 ottobre 2015"
output: html_document
---

## Requirements
The following R packages are needed:
```{r message=FALSE}
require(lubridate)
require(dplyr)
require(ggplot2)
options(scipen = 1)
```

## Loading and preprocessing the data
It is supposed that activity.csv is in the same folder of this Rmd file.
Data from activity.csv are loaded and dates are properly converted.
```{r}
activity<-read.csv("activity.csv",stringsAsFactors = FALSE)
activity$date<-ymd(activity$date)
```

## What is mean total number of steps taken per day?
In order to answer to this question, activity was grouped by date and it was calculated the total number of steps taken per day ignoring NA values.
```{r}
activity_by_date<-group_by(activity,date)
tot_steps_by_date<-summarise(activity_by_date,number=sum(steps,na.rm=TRUE))
```
The following figure shows the distribution of the total number of steps taken each day
```{r fig.width=5, fig.height=5, message=FALSE }
ggplot(tot_steps_by_date, aes(x=number))+
        geom_histogram()+
        labs(title="Distribution of total number of steps",
             x="Total number of steps",
             y="Number of occurrence")+
        theme_bw()
```

The following table summarize the mean and the median (rounded) of the previous distribution

Stat. Quantity  | Value
 ------------- | -------------
 Mean          | `r round(mean(tot_steps_by_date$number),1)` 
 Median        | `r round(median(tot_steps_by_date$number),1)`
